\section{Ohai}

Ohai detects data about your operating system. It can be used standalone, but its primary purpose is to provide node data to Chef.

When invoked, it collects detailed, extensible information about the machine it's running on, including Chef configuration, hostname, FQDN, networking, memory, CPU, platform, and kernel data.

When Chef configures the node object during each Chef run, these attributes are used by the chef-client to ensure that certain properties remain unchanged. These properties are also referred to as automatic attributes (which, as you remember, impossible to override by attributes from cookbooks, environments, roles and nodes). For example:

\begin{lstlisting}[language=Bash,label=lst:cookbook-ohai1]
node['platform'] # The platform on which a node is running. This attribute helps determine which providers will be used.
node['platform_version']  # The version of the platform. This attribute helps determine which providers will be used.
node['hostname']  # The host name for the node.
\end{lstlisting}

\subsection{Example}

\subsubsection{Recipe node}

Let's create new recipe <<node>>, which will install \href{http://nodejs.org/}{Node.js} on nodes in our \inline{my_cool_app} cookbook.

New default attributes:

\lstinputlisting[label=lst:cookbook-ohai-attributes,title=my-server-cloud/site-cookbooks/my\_cool\_app/attributes/node.rb]{code/book/cookbook/ohai/attributes.rb}

And recipe:

\lstinputlisting[label=lst:cookbook-ohai-recipe-before,title=my-server-cloud/site-cookbooks/my\_cool\_app/recipes/node.rb]{code/book/cookbook/ohai/node_before.rb}

As you can see, we using \inline{node['platform_family']} Ohai variable, which help for us understand type of OS on node and install needed package.

Also we used cookbook \inline{build-essential}, because we will install node.js from source. In this case we should add it as dependency in our metadata.rb file:

\begin{lstlisting}[label=lst:cookbook-ohai-meta]
depends 'nginx',           '~> 2.2.0'
depends 'build-essential'
\end{lstlisting}

To use this recipe we should add it in \inline{run_list}:

\begin{lstlisting}[language=JSON,label=lst:cookbook-ohai2]
{
  "run_list": [
    "recipe[my_cool_app]",
    "recipe[my_cool_app::node]"
  ]
}
\end{lstlisting}

But I want run this recipe with default recipe, so I will leave \inline{run_list} as is and add node recipe in default:

\begin{lstlisting}[label=lst:cookbook-ohai3]
...
include_recipe 'my_cool_app::node'
\end{lstlisting}

And check how it works:

\begin{lstlisting}[language=Bash,label=lst:cookbook-ohai4]
$ knife cookbook upload my_cool_app
Uploading my_cool_app    [0.1.0]
Uploaded 1 cookbook.

$ vagrant provision chef_second_client
INFO: remote_file[/usr/local/src/node-v0.10.26.tar.gz] created file /usr/local/src/node-v0.10.26.tar.gz
INFO: remote_file[/usr/local/src/node-v0.10.26.tar.gz] updated file contents /usr/local/src/node-v0.10.26.tar.gz
INFO: remote_file[/usr/local/src/node-v0.10.26.tar.gz] mode changed to 644
INFO: execute[tar --no-same-owner -zxf node-v0.10.26.tar.gz] ran successfully
INFO: bash[compile node.js] ran successfully
INFO: execute[nodejs make install] ran successfully

$ vagrant ssh chef_second_client

vagrant@precise64:~$ node -v
v0.10.26
\end{lstlisting}

As you can see it works.

\subsubsection{Ohai plugin}

In our node recipe we used maybe not best way to check node version, which already installed on node (if version mistmatch - we should install needed). Let's create Ohai plugin, which will give for use node.js version from server. First of all create in \inline{my_cool_app} new recipe \inline{ohai_plugin.rb} with content:

\lstinputlisting[label=lst:cookbook-ohai-recipe1,title=my-server-cloud/site-cookbooks/my\_cool\_app/recipe/ohai\_plugin.rb]{code/book/cookbook/ohai/ohai_plugin.rb}

This recipe will generate ohai plugin from \inline{system_node_js.rb.erb} template. Next we should create this template in folder <<templates/default/plugins>>:

\lstinputlisting[label=lst:cookbook-ohai-template1,title=my-server-cloud/site-cookbooks/my\_cool\_app/templates/default/plugin/system\_node\_js.rb.erb]{code/book/cookbook/ohai/system_node_js.rb.erb}

In first two lines we set by method \inline{provides} automatic attributes, which will provide for us this plugin.
Most of the information we want to lookup would be nested in some way, and ohai tends to do this by storing the data in a Mash. This can be done by creating a new mash and setting the attribute to it. We did this with \inline{system_node_js}. In the end of code, plugin set the version of node.js, if node.js installed on node. That's it!

Also we should add new dependency for our cookbook - ohai cookbook:

\begin{lstlisting}[label=lst:cookbook-ohai5]
depends 'nginx',           '~> 2.2.0'
depends 'build-essential'
depends 'ohai'
\end{lstlisting}

Next, let's try this plugin by adding node.rb recipe this content:

\begin{lstlisting}[label=lst:cookbook-ohai6]
include_recipe "build-essential"

include_recipe "my_cool_app::ohai_plugin"
Chef::Log.info "Installed Node version: #{node['system_node_js']['version']}" if node['system_node_js']

case node['platform_family']
  when 'rhel','fedora'
    package "openssl-devel"
  when 'debian'
    package "libssl-dev"
end
\end{lstlisting}

In this case we can little change our node.js recipe:

\begin{lstlisting}[label=lst:cookbook-ohai7]
execute "nodejs make install" do
  environment({"PATH" => "/usr/local/bin:/usr/bin:/bin:$PATH"})
  command "make install"
  cwd "/usr/local/src/node-v#{node['my_cool_app']['nodejs']['version']}"
  not_if {node['system_node_js'] && node['system_node_js']['version'] == node['my_cool_app']['nodejs']['version'] }
end
\end{lstlisting}

Now we can check how it works:

\begin{lstlisting}[language=Bash,label=lst:cookbook-ohai8]
$ knife cookbook upload my_cool_app
Uploading my_cool_app    [0.1.0]
Uploaded 1 cookbook.

$ vagrant provision chef_second_client
...
INFO: template[/etc/chef/ohai_plugins/system_node_js.rb] created file /etc/chef/ohai_plugins/system_node_js.rb
INFO: template[/etc/chef/ohai_plugins/system_node_js.rb] updated file contents /etc/chef/ohai_plugins/system_node_js.rb
INFO: template[/etc/chef/ohai_plugins/system_node_js.rb] owner changed to 0
INFO: template[/etc/chef/ohai_plugins/system_node_js.rb] group changed to 0
INFO: template[/etc/chef/ohai_plugins/system_node_js.rb] mode changed to 755
...
$ vagrant provision chef_second_client
...
WARN: Current  service[nginx]: /var/chef/cache/cookbooks/nginx/recipes/source.rb:123:in `from_file'
INFO: Installed Node version: 0.10.26
WARN: Cloning resource attributes for package[libssl-dev] from prior resource (CHEF-3694)
...
\end{lstlisting}

As you can see, after first \inline{provision} (\inline{knife cook}, \inline{knife ssh 'sudo chef-client'}) execution Ohai plugin will be installed, but not executed. Only on second execution chef-client will use Ohai plugin (because Ohai plugins load before kitchen).
